### 2020-05-15 paper/manuscript.Rmd
EVW comments

---
title: Interisland movements and oceanic swimming of woodland caribou in Newfoundland
author: Quinn M.R. Webber^1\*^, Jack G. Hendrix^1^, Alec L. Robitaille^2^,  Eric Vander Wal^1,2^
date: 
    ^1^Cognitive and Behavioural Ecology Interdisciplinary Program, Memorial
   University of Newfoundland
    ^2^Department of Biology, Memorial University of Newfoundland, Canada
    #\*Corresponding author: Quinn M.R. Webber (webber.quinn@gmail.com)
output:
   github_document: default
   bookdown::word_document2:
     reference_docx: template.docx
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
bibliography: ../references/swim.bib
csl: ecology.csl
---

```{r knitropts, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      fig.align = 'center')
```

```{r pkgs}
library(data.table)
```

```{r input}
edges <- readRDS('../output/island-edges.Rds')
icefree <- edges[season == 'icefree']
```

```{r summary-stats}
## Only icefree
# Number of events
nevents <- icefree[, .N]

# Number of individuals
nids <- icefree[, uniqueN(ANIMAL_ID)]

# Number of events by individual
neventid <- icefree[, .N, ANIMAL_ID]

# Max per individual
neventidMax <- max(neventid$N)

# Mean per individual
neventidMean <- round(mean(neventid$N))

# Min per individual
neventidMin <- min(neventid$N)

# IDs > 10 swimming
n10 <- length(icefree[, .N, ANIMAL_ID][N > 10]$N)

# Duration of time per island
dur <- round(icefree[, mean(islandlen)], digits = 0)
durMin <- round(icefree[, min(islandlen)], digits = 0)
durMax <- round(icefree[, max(islandlen)], digits = 0)

# Regular swimmers
ind <- edges[ANIMAL_ID == "FO2016011" | ANIMAL_ID == "FO2017001" | ANIMAL_ID == "FO2017013"]
indMean <- round(ind[, mean(islandlen)], digits = 0)

# Irregular swimmers
ind2 <- edges[ANIMAL_ID != "FO2016011" | ANIMAL_ID != "FO2017001" | ANIMAL_ID != "FO2017013"]
ind2Mean <- round(ind2[, mean(islandlen)], digits = 0)

## Full
# Number of events by individual + icefree season
nfull <- edges[, .N, .(ANIMAL_ID, season)]
```

\newpage

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between two small islands (Figure
1, see Figure 2 for location), a distance of at least 470 m which took ~9
minutes (~52 m per minute). Residents of Fogo Island confirmed this phenomenon
and have also reported regularly observing caribou swimming between islands.
This observation prompted us to investigate whether swimming behaviour was
common among GPS radio-collared caribou in the population [n = 29, for collaring
details see, @Peignier2019]. We identified an individual's swimming events as
two consecutive GPS locations (2-hour relocation rates)  on different islands.
Northeastern Newfoundland typically experiences pack ice during winter and
caribou are known locally to travel between islands by walking over the ice. We
restricted our GPS data to the ice-free period of the year (April 1 to December
31) and identified `r nevents` swimming events over three years (Figure 2).

<!-- That first paragraph would benefit from a bit of signpost at its end to
let us know what kind of ideas you're going go talk about near the end. -->

In total, 12 of 29 collared female caribou swam between islands in the Fogo
Island archipelago. Of the caribou that swam, caribou swam on average
approximately `r neventidMean` (range = `r neventidMin`-`r neventidMax`) times
per year and remained on each island for approximately `r dur` (range = 
`r durMin`-`r durMax`) days before swimming again. Based on GPS data to infer
swimming, we separated the population into three apparent groups of individuals:
those that swam regularly between islands (n = 3, every ~`r indMean` days);
those that rarely engaged in swimming events (n = 9, every ~ `r ind2Mean` days);
and those that never swam (n = 17).

Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond2016],
to avoid predators [@Bergerud1985], and to access islands during calving
[@Bergerud1990]. Even for adept swimmers like caribou, which have hollow hair
shafts that aid in flotation, the energetic expenditure of swimming for
quadrupedal mammals is significantly higher than walking or running, and
drowning is also possible [@Miller1985]. Despite the abundance of coastal and
island caribou herds, only a few studies have documented caribou swimming
in the ocean, and most of these lack observation, but rather deduce swimming
based on presence of caribou on a previously uninhabitated island  [e.g.
@Miller1995; @Miller2002; @Jeffery2007; @Ricca2012]. Caribou may swim between
islands in the ocean for similar reasons they swim in freshwater [@Leblond2016],
mainly concerning predator avoidance or to access new habitats and foraging
opportunities. We propose that one such explanation, the forage limitation
hypothesis, is the most likely explanation for oceanic swimming for caribou that
live on islands in the Fogo Island archipelago.

Caribou were introduced to Fogo Island (n = 26 animals between 1964–67) and
Change Island (n = 5 animals in 1964) from Newfoundland as part of a series of
translocations and introductions throughout the province [@Bergerud1989].
Typical caribou predators such as black bears (*Ursus americanus*) and wolves
(*Canis lupus*) are absent from Fogo Island; the dominant predator is coyote
(*Canis latrans*), which predate calves but it remains unclear what threat they
pose, if any, to adult caribou [@Rayl2014]. Given the minimal risk of predation
to adults, we propose that forage limitation is more likely to drive movement
between islands than predator avoidance. For several decades after their
introduction to Fogo Island, the caribou population did not exceed ~100
individuals [@Bergerud1989, Newfoundland and Labrador Wildlife Division,
unpublished data]. During the 1990s, population density reached ~300 animals and
anecdotal evidence suggests caribou were first seen on nearby islands during
this same period. Increased competition for resources on Fogo Island due to this
increase in density could have precipitated movement to other islands.

Forage limitation and over-grazing is a major concern for caribou populations
and can lead to reduced female reproductive success [@Schaefer2016]. Forage
scarcity has been proposed as a potential reason for caribou moving between
arctic islands on the sea-ice in winter [@Miller1977]. For the Fogo Island
population, as the population grew and predation pressure was insufficient to
maintain population density, forage depletion could lead caribou to make
similar, but riskier and more costly, movements between islands during the
ice-free season. Given the small size of other islands near Fogo (Figure 2),
over-grazing by newly arrived caribou could rapidly deplete forage on an annual
cycle [@Bergerud1990], leading to more frequent movements back and forth between
islands rather than permanent relocation.

The trade-off associated with swimming between islands can be understood as
density-dependent habitat selection governed by the Ideal Free Distribution
[@Morris1987; @Bradbury2015]. Ideal Free Distribution theory predicts a fitness
equilibrium: when the density in a given habitat patch has exceeded the optimum
for fitness within that patch, animals should relocate and settle new habitat
patches, so that fitness is equal across all patches [@Bradbury2015]. Islands
act as discrete habitat patches in this case, but the costs associated with
swimming create a trade-off when moving between patches. For caribou in the Fogo
Island archipelago, the role of forage limitation in habitat selection patterns
remains unknown, but swimming between islands may be at least partially governed
by density-dependent habitat selection.

Forage limitation may be the ultimate explanation for caribou swimming between
oceanic islands, but a number of proximate mechanisms remain untested. Our
observations of caribou swimming in the ocean lead to five questions for future
research:
<!-- a) For the questions that arise, maybe try to generalize the language a bit
so that they are a bit broader: caribou as model, but emphasize the
ecological/evolutionary process it will comment on. 
b.) Consider tweaking the
intro the the questions less as 'someone could do this' as something along the
lines of : this has ignited in us the idea that we can tailor our data
collection in our closely studies system to access some of these questions. -->

1) How do the energetic cost and risk of swimming differ from walking on sea
ice? Walking over solid ice is likely less energetically expensive than
swimming, but may pose a greater risk of falling through depending on the
thickness and stability of ice. What proportion of inter-island movements are
swimming events compared to walking on ice? Although the presence and extent of sea ice
varies annually,  we
deduce that although the majority of inter-island movements occur between April
1 and December 31, some inter-island movements appear to be caribou walking on
sea ice between islands (Figure 2B). Future research on the trade-offs of these types of
movement should consider both how the costs associated with each differ, but
also how the potential benefits - foraging opportunities - may vary seasonally.

2) At which temporal and spatial scales do caribou choose to swim
between islands? Does the time of day, weather, season, or availability
of food influence swimming?

3) Does reproductive status (presence of a calf-at-heel) or demographic class
(age or sex) influence propensity for oceanic swimming? Our initial observation
was of an adult male, but all collared individuals were adult females. Are
younger caribou more likely to swim if they are unable to compete with older
individuals for resources? Are males or females without a calf-at-heel more
likely to swim in the summer given they are not restricted based on presence of
a calf-at-heel, as females are?

4) If swimming is a function of density-dependent habitat selection, what are
the costs of movement between patches? IFD assumes zero-cost movement, but could
the risk and energetic costs associated with oceanic swimming be estimated using
the potential fitness deficit that caribou are willing to tolerate before
swimming to a new island?

5) In relation to all the previous questions: how do individuals vary in their
assessment of the costs and benefits of swimming between islands? Why did some
caribou regularly engage in swimming events, while more than half of our marked
population were never observed swimming in the ocean? The forage limitation
hypothesis is thought to apply uniformly across populations, but individual
caribou could evaluate the costs and benefits of swimming differently. It is
possible that some individuals, in our case ~10%, considered swimming to be less
costly or perceived competition or predation more acutely than their
conspecifics.

We surmise that for Fogo Island, as competition among conspecifics increased
along with population size, density-dependent habitat selection resulted in
expansion of the population to nearby islands. Following Ideal Free Distribution
theory, caribou should swim to new islands when the average fitness of
individuals on the starting island exceeds the density-fitness equilibrium
[@Morris1987]. While this is an ultimate explanation, our data suggest caribou
only remain on smaller islands for a few days at a time, so the fitness
equilibrium remains a theoretical construct. More likely, however, is that
fine-scale competition, density-dependent habitat selection, and forage
depletion drives individuals to periodically swim between islands. Our focal
observations add to the evidence that caribou can, and occasionally do, swim in
the ocean [e.g. @Miller1995; @Miller2002;
@Jeffery2007; @Ricca2012]. We suggest that swimming is likely more common than
previously thought for caribou living on oceanic islands, and that forage
limitation and the associated density-dependent habitat selection is an ultimate
explanation for this phenomenon. Although our inference is limited to
observations, islands appear to represent discrete foraging patches for
terrestrial animals that can influence fitness via increased foraging
opportunities. We also present several hypotheses related to the nuances and
further variation within the idea that density-dependent habitat selection
governs swimming behaviour, and posit that above and beyond these additional
factors, individuals can vary in their evaluation of costs and benefits of
movement between these habitat patches.

**Acknowledgements**
We thank M. Laforge, M. Bonar, C. Hart, and R. Huang for help in the field.
Logistical support was provided by L. Bixby. We thank all members of the
Wildlife Evolutionary Ecology Lab, including C. Hart, C. Prokopenko, J. Kennah,
J.W. Turner, and S. Boyle for their comments on previous versions of this
manuscript. Newfoundland Wildlife Division including S. Moores, B. Adams, C.
Doucet, W. Barney, and J. Neville for logistical support in the field as well as
help with data data collection and management. We thank T. Bergerud and S.
Mahoney for their vision in initiating much of the work on caribou in
Newfoundland. Funding for this study was provided by the National Sciences and
Engineering Research Council (QMRW, JGH, EVW). We respectfully acknowledge the
territory in which data were collected and analyzed as the ancestral homelands
of the Beothuk, and the island of Newfoundland as the ancestral homelands of the
Mi’kmaq and Beothuk.

\newpage

(ref:fig1) Photograph of swimming unmarked adult male caribou (*Rangifer tarandus*) from Western to Eastern Indian Island taken on 30 May 2017.

```{r, fig.cap = '(ref:fig1)', fig.width=4}
knitr::include_graphics('../graphics/IMG_1562_crop_edit.jpg')
```

\newpage

(ref:fig2) A) Map of the Fogo Island archipelago with swimming events between islands. B) Histogram displaying the frequency of swimming events throughout the year. Note, colours correspond to individual caribou and vertical black bars delineate the ice-free season (April 1 to December 31). C) Inset of swimming events between small islands on the northern coast of Fogo island. D) Inset of swimming events between islands on the southern coast of Fogo island. Note, the grey diamond represents the location we observed an adult male caribou swimming (see Figure 1).

```{r, fig.cap = '(ref:fig2)', fig.width=4}
knitr::include_graphics('../graphics/Fig2.png')
```

<!-- edit caption Fogo Island is the largest
island in the archipelago (237.71 km^2^), but there are at least three other
large islands: Western Indian (77.6 km^2^), Eastern Indian (38.7 km^2^) and
Change (XX km^2^) Islands, as well as numerous smaller islands, including
Blundon’s (1.18 km^2^), North Long (1.01 km^2^), South Long (0.48 km^2^), Kate’s
(1.64 km^2^), and Brother’s (1.59 km^2^) Islands. Habitats are similar across
the archipelago, consisting largely of coniferous and mixed forests of balsam
fir (*Abies balsamea*), black spruce (*Picea mariana*), and white birch 
(*Betula papyrifera*) as well as bogs, lakes, lichen and rocky
barrens.-->

\newpage

**Literature Cited**