### 2021-04-12 paper/manuscript.Rmd
new P2

---
title: On the marginal value of swimming in woodland caribou
author:
  - name: Quinn M.R. Webber
    institute: cabe
    correspondence: true
  - name: Jack G. Hendrix
    institute: cabe
  - name: Alec L. Robitaille
    email: robit.alec@gmail.com
    institute: bio
  - name: Eric Vander Wal
    institute: [bio, cabe]
institute: 
  - bio: Department of Biology, Memorial University of Newfoundland, Canada
  - cabe: Cognitive and Behavioural Ecology Interdisciplinary Program, Memorial University of Newfoundland 
output:
   bookdown::word_document2:
     reference_docx: template.docx
     pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: ../references/swim.bib
csl: ecology.csl
---

```{r knitropts, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      fig.align = 'center')
```

```{r pkgs}
library(data.table)
```

```{r input}
edges <- readRDS('../output/island-edges.Rds')
icefree <- edges[season == 'icefree']
```

```{r summary-stats}
## Only icefree
# Number of events
nevents <- icefree[, .N]

# Number of individuals
nids <- icefree[, uniqueN(ANIMAL_ID)]

# Number of events by individual
neventid <- icefree[, .N, ANIMAL_ID]

# Max per individual
neventidMax <- max(neventid$N)

# Mean per individual
neventidMedian <- round(median(neventid$N))

# Min per individual
neventidMin <- min(neventid$N)

# IDs > 10 swimming
n10 <- length(icefree[, .N, ANIMAL_ID][N > 10]$N)

# Duration of time per island
dur <- round(icefree[, mean(islandlen)], digits = 0)
durMin <- round(icefree[, min(islandlen)], digits = 0)
durMax <- round(icefree[, max(islandlen)], digits = 0)

# Regular swimmers
ind <- edges[ANIMAL_ID == "FO2016011" | ANIMAL_ID == "FO2017001" | ANIMAL_ID == "FO2017013"]
indMean <- round(ind[, mean(islandlen)], digits = 0)

# Irregular swimmers
ind2 <- edges[ANIMAL_ID != "FO2016011" | ANIMAL_ID != "FO2017001" | ANIMAL_ID != "FO2017013"]
ind2Mean <- round(ind2[, mean(islandlen)], digits = 0)

## Full
# Number of events by individual + icefree season
nfull <- edges[, .N, .(ANIMAL_ID, season)]
```

\newpage

Movement is costly. Even for adept swimmers like caribou (Rangifer tarandus), which have hollow hair shafts that aid in flotation, the energetic expenditure of swimming for terrestrial mammals is markedly higher than walking or running (Miller and Gunn 1985). The drivers that promote the decision to swim for terrestrial mammals, like caribou, are often related to the profitability of forage (Miller et al. 1977) and the risks of predation (Jeffery et al. 2007, Jordan et al. 2010). The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to approximately 300 woodland caribou, descendants of 26 individuals introduced from the Island of Newfoundland to Fogo Island between 1964–67 (Bergerud and Mercer 1989). During fieldwork on May 30, 2017, we observed an unmarked adult male caribou swim between two smaller islands (Figure 1) a distance of 470 m which took approximately 9 minutes. Given that swimming is energetically expensive, we ask how often do caribou swim between islands, how long do they occupy islands, and ultimately what motivates swimming? Here, we frame swimming and island residency as a model of patch use as described by the marginal value theorem (Charnov 1976) within the context of the forage limitation hypothesis. For caribou, the forage limitation hypothesis is the constraints on food availability associated with changes in abundance through time (Klein 1991, Schaefer et al. 2016). Here, we integrate our understanding of forage limitation in caribou populations by using island residency as a proxy for time in patch and island size as a proxy for energy intake in a patch (Figure 2).

Forage limitation and over-grazing is a concern for caribou populations and can lead to reduced female reproductive success (Klein 1991, Schaefer et al. 2016). Forage scarcity has been proposed as a potential reason for caribou moving between arctic islands on the sea-ice in winter (Miller et al. 1977). We submit that testing the marginal value theorem could improve our understanding of forage limitation. The marginal value theorem posits that animals will forage optimally by depleting the resources in a given patch (i.e. island) before moving to a new patch (Charnov 1976). Depletion of resources on small islands could therefore mirror patterns of over-grazing and forage limitation that occur for caribou herds at higher densities in Newfoundland (Schaefer et al. 2016) and elsewhere in their range (Klein 1991).

`r nevents` 
In total, 12 of 29 collared female caribou swam between islands in the Fogo
Island archipelago. Of the caribou that swam, caribou swam on average
approximately `r neventidMean` (range = `r neventidMin`-`r neventidMax`) times
per year and remained on each island for approximately `r dur` (range = 
`r durMin`-`r durMax`) days before swimming again. Based on GPS data to infer
swimming, we separated the population into three apparent groups of individuals:
those that swam regularly between islands (n = 3, every ~`r indMean` days);
those that rarely engaged in swimming events (n = 9, every ~ `r ind2Mean` days);
and those that never swam (n = 17).

Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond2016],
to avoid predators [@Bergerud1985], and to access islands during calving
[@Bergerud1990]. Even for adept swimmers like caribou, which have hollow hair
shafts that aid in flotation, the energetic expenditure of swimming for
quadrupedal mammals is significantly higher than walking or running, and
drowning is also possible [@Miller1985]. Despite the abundance of coastal and
island caribou herds, only a few studies have documented caribou swimming
in the ocean, and most of these lack observation, but rather deduce swimming
based on presence of caribou on a previously uninhabitated island  [e.g.
@Miller1995; @Miller2002; @Jeffery2007; @Ricca2012]. Caribou may swim between
islands in the ocean for similar reasons they swim in freshwater [@Leblond2016],
mainly concerning predator avoidance or to access new habitats and foraging
opportunities. We propose that one such explanation, the forage limitation
hypothesis, is the most likely explanation for oceanic swimming for caribou that
live on islands in the Fogo Island archipelago.

Caribou were introduced to Fogo Island between 1964–67 as part of a series of
translocations throughout the province [@Bergerud1989]. Typical caribou
predators such as black bears (*Ursus americanus*) and wolves (*Canis lupus*)
are absent from Fogo Island; the dominant predator is coyote (*Canis latrans*),
which predate calves but it remains unclear what threat they pose, if any, to
adult caribou [@Rayl2014]. Given the minimal risk of predation to adults and low
density of coyote on Fogo Island, we propose that forage limitation is more
likely to drive movement between islands than predator avoidance. For several
decades after their introduction to Fogo Island, the caribou population did not
exceed ~100 individuals [@Bergerud1989, Newfoundland and Labrador Wildlife
Division, unpublished data]. During the 1990s, population density reached ~300
animals and anecdotal evidence suggests caribou were first seen on nearby
islands during this same period. Increased competition for resources on Fogo
Island due to this increase in density could have precipitated movement to other
islands.

Forage limitation and over-grazing is a major concern for caribou populations
and can lead to reduced female reproductive success [@Schaefer2016]. Forage
scarcity has been proposed as a potential reason for caribou moving between
arctic islands on the sea-ice in winter [@Miller1977]. For the Fogo Island
population, as the population grew and predation pressure was insufficient to
maintain population density, forage depletion could lead caribou to make
similar, but riskier and more costly, movements between islands during the
ice-free season. Given the small size of other islands near Fogo (Figure 2),
over-grazing by newly arrived caribou could rapidly deplete forage on an annual
cycle [@Bergerud1990], leading to more frequent movements back and forth between
islands rather than permanent relocation.

The trade-off associated with swimming between islands can be understood as
density-dependent habitat selection governed by the Ideal Free Distribution
[@Morris1987]. Ideal Free Distribution theory predicts a fitness
equilibrium: when the density in a given habitat patch has exceeded the optimum
for fitness within that patch, animals should relocate and settle new habitat
patches, so that fitness is equal across all patches [@Morris1987]. Islands
act as discrete habitat patches in this case, but the costs associated with
swimming create a trade-off when moving between patches. For caribou in the Fogo
Island archipelago, the role of forage limitation in habitat selection patterns
remains unknown, but swimming between islands may be at least partially governed
by density-dependent habitat selection.

Forage limitation may be the ultimate explanation for caribou swimming between
oceanic islands, but a number of proximate mechanisms remain untested. Our
observations of caribou swimming in the ocean lead to five questions for future
research:
<!-- a) For the questions that arise, maybe try to generalize the language a bit
so that they are a bit broader: caribou as model, but emphasize the
ecological/evolutionary process it will comment on. 
b.) Consider tweaking the
intro the the questions less as 'someone could do this' as something along the
lines of : this has ignited in us the idea that we can tailor our data
collection in our closely studies system to access some of these questions. -->

1) How do energetic costs and risks of swimming differ from walking on sea ice?
Walking over ice is less energetically expensive than swimming, but may pose a
greater risk of mortality from falling through the ice. Although the presence
and extent of sea ice varies annually,  we deduce that although the majority of
inter-island movements occur between April 1 and December 31, some inter-island
movements appear to be caribou walking on sea ice between islands (Figure 2B).
Our aim for future research on the trade-offs of these types of movement
considers how the costs, e.g. energy expenditure or mortality, and benefits,
e.g. novel foraging opportunities, may vary seasonally.

<!--2) At which temporal and spatial scales do caribou choose to swim
between islands? Does the time of day, weather, season, or availability
of food influence swimming?-->

<!--3) Does reproductive status (presence of a calf-at-heel) or demographic class
(age or sex) influence propensity for oceanic swimming? Our initial observation
was of an adult male, but all collared individuals were adult females. Are
younger caribou more likely to swim if they are unable to compete with older
individuals for resources? Are males or females without a calf-at-heel more
likely to swim in the summer given they are not restricted based on presence of
a calf-at-heel, as females are? -->

2) If swimming is a function of density-dependent habitat selection, what are
the costs of movement between patches? IFD assumes zero-cost movement, but could
the risk and energetic costs associated with oceanic swimming be estimated using
the potential fitness deficit that caribou are willing to tolerate before
swimming to a new island?

3) In relation to questions 1 and 2: how do individuals vary in their
propensity to swim and their assessment of the costs and benefits of swimming
between islands? We observed some individuals that regularly engaged in swimming
events, while others were never observed swimming in the ocean? Our existing
research on individual behavioural variation in caribou movement ecology
]@Webber2020] provides a basis to investigate the role of inherent individual
differences in swimming behaviour. Specifically, the forage limitation
hypothesis applies uniformly across populations, but individuals could evaluate
the costs and benefits of swimming differently.

We surmise that for Fogo Island, as competition among conspecifics increased
along with population size, density-dependent habitat selection resulted in
expansion of the population to nearby islands. Following Ideal Free Distribution
theory, caribou should swim to new islands when the average fitness of
individuals on the starting island exceeds the density-fitness equilibrium
[@Morris1987]. While this is an ultimate explanation, our data suggest caribou
only remain on smaller islands for a few days at a time, so the fitness
equilibrium remains a theoretical construct. More likely, however, is that
fine-scale competition, density-dependent habitat selection, and forage
depletion drives individuals to periodically swim between islands. Our focal
observations add to the evidence that caribou can, and occasionally do, swim in
the ocean [e.g. @Miller1995; @Miller2002;
@Jeffery2007; @Ricca2012]. We suggest that swimming is likely more common than
previously thought for caribou living on oceanic islands, and that forage
limitation and the associated density-dependent habitat selection is an ultimate
explanation for this phenomenon. Although our inference is limited to
observations, islands appear to represent discrete foraging patches for
terrestrial animals that can influence fitness via increased foraging
opportunities. We also present several hypotheses related to the nuances and
further variation within the idea that density-dependent habitat selection
governs swimming behaviour, and posit that above and beyond these additional
factors, individuals can vary in their evaluation of costs and benefits of
movement between these habitat patches.

**Acknowledgements**
We thank M. Laforge, M. Bonar, C. Hart, and R. Huang for help in the field.
Logistical support was provided by L. Bixby. We thank all members of the
Wildlife Evolutionary Ecology Lab, including C. Hart, C. Prokopenko, J. Kennah,
J.W. Turner, and S. Boyle for their comments on previous versions of this
manuscript. Newfoundland Wildlife Division including S. Moores, B. Adams, C.
Doucet, W. Barney, and J. Neville for logistical support in the field as well as
help with data data collection and management. We thank T. Bergerud and S.
Mahoney for their vision in initiating much of the work on caribou in
Newfoundland. Funding for this study was provided by the National Sciences and
Engineering Research Council (QMRW, JGH, EVW). We respectfully acknowledge the
territory in which data were collected and analyzed as the ancestral homelands
of the Beothuk, and the island of Newfoundland as the ancestral homelands of the
Mi’kmaq and Beothuk.

\newpage

(ref:fig1) Photograph of swimming unmarked adult male caribou (*Rangifer tarandus*) from Western to Eastern Indian Island taken on 30 May 2017.

```{r, fig.cap = '(ref:fig1)', fig.width=4}
knitr::include_graphics('../graphics/IMG_1562_crop_edit.jpg')
```

\newpage

(ref:fig2) A) Map of the Fogo Island archipelago with swimming events between islands. B) Histogram displaying the frequency of swimming events throughout the year. Note, colours correspond to individual caribou and vertical black bars delineate the ice-free season (April 1 to December 31). C) Inset of swimming events between small islands on the northern coast of Fogo island. D) Inset of swimming events between islands on the southern coast of Fogo island. Note, the grey diamond represents the location we observed an adult male caribou swimming (see Figure 1).

```{r, fig.cap = '(ref:fig2)', fig.width=4}
knitr::include_graphics('../graphics/Fig2.png')
```

<!-- edit caption Fogo Island is the largest
island in the archipelago (237.71 km^2^), but there are at least three other
large islands: Western Indian (77.6 km^2^), Eastern Indian (38.7 km^2^) and
Change (XX km^2^) Islands, as well as numerous smaller islands, including
Blundon’s (1.18 km^2^), North Long (1.01 km^2^), South Long (0.48 km^2^), Kate’s
(1.64 km^2^), and Brother’s (1.59 km^2^) Islands. Habitats are similar across
the archipelago, consisting largely of coniferous and mixed forests of balsam
fir (*Abies balsamea*), black spruce (*Picea mariana*), and white birch 
(*Betula papyrifera*) as well as bogs, lakes, lichen and rocky
barrens.-->

\newpage

```{r tab1}
knitr::kable(readRDS('../output/table1.Rds'),
             caption = '')
```

\newpage

**Literature Cited**